TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2018-08-13 paper/draft.tex
intro text

Why \thecannon\ approach? brief importance statment cannon and what it does (cite \citealt{Ness:2015}, \citealt{Ho:2017a}, \citealt{Casey:2016})

\begin{enumerate}
\item This paper will refer to any physical or empirical measurement of a star as \emph{stellar label}

\item Can be used to amplify a small number of labeled M-dwarfs in a training set into a huge number of labeled M-dwarfs, provided that there are overlapping objects;

\item Make the point that \thecannon\ does not require us to have good M-dwarf models in the \apogee\ wavelength range at all

\item better than template matching in that it is using less sparsely sampled known mdwarfs
\end{enumerate}

%photometric data, HR digram, comparison to evolutionary models

>>>>

In this work we build a data-driven model for M dwarf spectra with \thecannon (\citealt{Ness:2015}, \citealt{Ho:2017a}, \citealt{Casey:2016}).
\thecannon is a regression (typically quadratic) for the spectral pixels given a training
set of stars with reliable or accurate or assumed-accurate labels.
The model predicts spectrum given labels, but it can be turned around and also
used to predict labels given spectrum.
Thus it is a generative model for stellar spectra that can be used for parameter inference.

The data-driven approach of \thecannon\ makes sense in various contexts:
One is the context in which stellar labels are only known for a small number of stars
but there are spectra taken for many more.
Another is the context in which it is very expensive to obtain labels for a star, but
there are many stars that need labels.
Another is the context in which there are spectral models that work in one wavelength
range or resolution but not in another.
The context of this paper is the latter:
There are working and reliable models of M dwarf stars in the optical, which have been
used successfully, but (as we note above) there isn't as much success with such modeling in the infrared,
and not at \apogee\ resolution.

Here we take M-dwarf labels from samples of optically observed stars that overlap the
\sdssiv\ \apogee\ sample, and use those labels to train a model and label all of the
\sdssiv\ \apogee\ M dwarfs.
One set of labels are physical parameters (effective temperatures and metallicities)
and the other set of labels are spectral types.
In each case, \thecannon\ produces a predictive model for \apogee\ spectra, but without
any input of stellar physics beyond whatever went in to producing the labels in the first
place, which was in the context of optical spectroscopy.
That is, the model to be applied to the \apogee\ data is learned from the \apogee\ data alone
This obviates physical models in the infrared, and capitalizes on the excellent data quality
delivered by the \apogee\ spectrograph and project.
